### 2019-02-19 Code/README.Rmd
Beginning move to standard project directory organization

---
title: "Code"
output: github_document
---

## Description

Each script should do something specific (like tidyverse functions), have a descriptive name, and include number prefixes if they are meant to be run in a certain order (e.g., `01_import_data.R`, `02_clean_data.R`).

## Project Workflow

* Use RStudio projects.
* Use `here::here()` to specify files paths.
* Use tidyverse functions as much as possible.
* Try and follow the [tidyverse style guide](https://style.tidyverse.org).