### 2019-02-19 Paper/README.Rmd

---
title: "Paper"
output: github_document
---

## Description

The manuscript of the paper, without any PDF knits.