### 2019-02-22 README.Rmd

---
title: "Assessing the Severity of Service Failures"
output: github_document
---

## Abstract

Firms employ a variety of passive listening platforms (e.g., customer comment cards, help lines, social media) that elicit a variety of complaints. Since these complaints are self-selected, it isn't obvious if complaints are widespread or genuinely indicative of systemic service failures. In other words, how can can a firm know when complaints can be ignored and when they need to addressed? This issue is related to a larger class of problems where the goal is to make inference about the size of a population from a non-random sample (e.g., the German tank problem). In this paper, we develop a model that uses a variety of observed customer complaints on Twitter to make inference about the severity of service failures.

For general details on GitHub usage, project organization, and project workflow, see [Research Assistant Training](https://github.com/marcdotson/ra-training).