### 2019-03-19 Paper/paper.Rmd

---
title: "Classifier"
output: github_document
---
# The Classification Model
A key component to successfuly using twitter data to assess service failures is the ability to classify tweets as complaints and non-complaints. If our team can accurately classify a tweet, we will be capable of tackling the greater issue of judging how serious the complaint is, and whether or not a company needs to take action.

While there are a massive number of methods to perform this task, our team has identified three tools to aid us in our classification: Principal Component Analysis (PCA), FastText, and XGBoost. Using these three tools, we've created a "frankenstein model" that incorporates all three tools listed above. The model has been trained on 2,500 tweets labeled 0 for "non-complaint" and 1 for "complaint." We have then tested the model on a similarly-sized dataset.

To use our model, we feed a set of tweets into a PCA model and a FastText classification model. The output of these two models is then fed into our XGBoost model, which makes a binary prediction for the tweet as a complaint or non-complaint.

Because our model uses three distinct modeling techniques, I will refer to each of these modeling techniques as "sub-models."

# The FastText Sub-Model
## Overview
In this section, we'll be documenting how we built our model for classifying tweets as "complaints" and "non-complaints." The code is almost entirely derived from the Code/cnn-classifier/R/fasttext_supervised_classifier.R script.

## Load Packages
```{r}
# library(fastrtext)
# library(tidyverse)
```

## Define Functions
No functions have been defined in this file.

## About fastText
fastText was designed by Facebook and it allows us to build models for both classification and word embeddings. Classification refers to analyzing a set of text, and labeling it according to the topics it closely relates to. Word embeddings help us predict which words are most closely related to a single word of interest.

fastText boasts that it can perform classification and the creation of word embeddings faster than other tools used for comparable activities (like Word2vec and GloVe) without sacrificing accuracy. This is what drew our team's attention to this method from the beginning.

If you are unfamiliar with Fasttext, we'd reccomend getting to know it better by following along with the tutorials found at https://fasttext.cc/docs/en/supervised-tutorial.html. The tutorial run entirely from the terminal (a.k.a. the command line, shell, bash, command prompt), so if you're uncomfortable navigating your computer from the terminal, it would be worthwhile to find some tutorials on navigating directories within your terminal before beginning to learn fastText. If you'd like to learn more about how fastText works, this post found on Medium is also helpful https://towardsdatascience.com/fasttext-under-the-hood-11efc57b2b3.

Finally, if you don't conceptually understand word embeddings and classification, we reccomend doing a bit of research to get a better grip on those ideas before proceding.

## Execute Code

### List File Paths of Model and Data
To build our fasttext classification model, we begin by defining where we want to store our fastText model once we've trained it. In our case, we want to store the model in Output/models and name the model class_mod_1.dat. Within R, we'll refer to this file path as file_model.

As a data input for training our model, we already have two text files: train_data_1.txt and test_data_1.txt. The file paths to these text files (already built) will be referred to as train_data and test_data.

```{r}
# file_model <- "Output/models/class_mod_1.dat"
# train_data_path <- "Output/clean_data/train_data_1.txt"
# test_data_path <- "Output/clean_data/test_data_1.txt"
```

### Read in the Data
We have our training and test data kept in one csv file. We'll read it in after we've filtered all short tweets (with greater than five characters). We'll name this object tweet_data.
```{r}
# This code (and all code below) is broken because we don't have access to a tweet_sample_5k_full.csv file.

# tweet_data <- read_csv("data/Marketing Research Labeled Tweets_ - tweet_sample_5k_FULL.csv") %>%  
#  filter(nchar(tweet_text) > 5)
```

### Format the Training Data
For any supervised learning model (such as complaint classification), our training and test datasets need to have data labeled with the "correct" answers. This will allow our model to "learn" the correct answers, and it will hopefully help our model accurately predict correct answers.

For our fastText model, we need to make sure each tweet is preceded by a label (in our case, the labels will be "complaint" and "non-complaint". fastText labels have two underscores before and after the label. For example, our complaint labels need to look like this: __label__complaint. Furthermore, the data we feed into the fastText model needs to be a character vector (like a .txt file), not a dataframe. We'll have to do some formatting before we get to that point.

Now that we have access to the data, we'll edit a variable named "label." Currently the labels are 0 for non-complaints and 1 for complaints. When we're done, we'll replace the 0's and 1's with their respective terms. Additionally, the label and text need to be in the same cell, so we'll make a new variable that concotenates the text and the label toghether.

Finally, we need to randomly and evenly tag our data: half of it will become training, and half of it will become test data.

```{r}
# test_train_data <- 
#   tweet_data %>% 
#   mutate(label = case_when(label == 0 ~ "non-complaint", # each line (tweet) is preceded by a label to be predicted
#                               label == 1 ~ "complaint"),
#             input_lines = paste(label, tweet_text)) %>% 
#   group_by(label) %>% # get a representative sample of both complaints and non complaints
#   mutate(test_train = c("test", "train")[rbinom(n(), 1, 0.8) + 1]) %>% # assign test train splits
#   ungroup()
```

#### Create the .txt Files
As mentioned earlier, we need to create a text file of our training and test datasets because the fastText model only takes character vectors as inputs. With our data labeled both for training/test and complaint/non-complaint, we're ready to do this. We will use the write_lines() function to do this. The train_data_path is the file location we specified earlier in this script.

!!!!!!!!!! I'm still not sure how we are using the test_labels_without_prefix. If its not the test data, what is it?

```{r}
# write train data
# train_data_lines <- 
#   test_train_data %>% 
#   filter(test_train == "train") %>% 
#   pull(input_lines) %>% 
#   paste0("__label__", .)
# 
# write_lines(train_data_lines, train_data_path)
# 
# # write test data
# test_data_lines <- 
#   test_train_data %>% 
#   filter(test_train == "test") %>% 
#   pull(input_lines) %>% 
#   paste0("__label__", .)
# 
# write_lines(test_data_lines, test_data_path)

# create a character vector containing the tweets to test without their labels
# test_labels_without_prefix <- 
#   test_train_data %>% 
#   filter(test_train == "test") %>% 
#   pull(label)
```

### Train the Model
The package we are using (fastrtext) uses the execute() function to train the model. This function makes more sense if you've followed the fastText tutorials, but put simply, it concotenates the arguments into a single command which is then run in the terminal.

After we train the model, we'll load the model into our environment.

"supervised" specifies that this is a supervised model (classification, as opposed to creating word vectors). 
"-input" specifies the file location of the training text files.
"-output" specifies the file location where we want to place our model, once it has finished training.

You can find more information about the parameters of the model (such as -dim, -lr, and -epoch) here: https://fasttext.cc/docs/en/options.html.

!!!!!! Considering that this is such a large dataset, it might be worthwhile to use -loss hs (hierarchical softmax) to speed up the training. It should not significantly affect accuracy.

```{r}
# train model
# execute(commands = c("supervised", "-input", train_data_path, "-output", file_model, "-dim", 20, "-lr", 1, "-epoch", 20, "-wordNgrams", 2, "-verbose", 1))
# 
# # load model
# model <- load_model(file_model)
```

### Review Model:
We'll want to see how a few of the predictions turned out. 
```{r}
# prediction are returned as a list with words and probabilities
# predictions <- predict(model, sentences = test_data_lines, simplify = TRUE, unlock_empty_predictions = TRUE)
# head(predictions, 5)
```

This will help us confirm that our model worked as expected.
```{r}
# length(predictions)
# length(test_labels_without_prefix)
# length(test_data_lines)
```

### Accuracy
Let's look at how accurately the model predicted names
```{r}
# # Compute accuracy
# mean(names(unlist(predictions)) == test_labels_without_prefix, na.rm = TRUE)
```

Here, we'll look at the hamming loss, and a few of the predictions.

The hamming loss is the number of wrong labels over the total number of labels. Hamming loss is most useful when applied to multi-classifiers because when it is done with binary classifiers (such as ours), it is simply equal to 1 - accuracy.
```{r}
# # because there is only one category by observation, hamming loss will be the same
# get_hamming_loss(as.list(test_labels_without_prefix), predictions)
# 
# # test predictions
# predictions <- predict(model, sentences = test_to_write)
# print(head(predictions, 5))
# 
# # you can get flat list of results when you are retrieving only one label per observation
# print(head(predict(model, sentences = test_to_write, simplify = TRUE)))
```

## Analysis

# The PCA Sub-Model
## Intro to PCA
The idea behind PCA is that a complex multi-dimensional problem can be boiled down to a single-dimensional problem. For example, what if you had a model that tried to predict a student's grade on a test by looking at how many hours the student studied, how many lectures the student attended, how many hours of sleep they got, what cereal they eat for breakfast, and how many siblings they have. While all of these factors could arguably have an affect on the student's test grade, I would expect the "hours studied" and "lectures attended" variables to have the strongest correlation to test. A PCA model would notice the most important variables of a model, and simplify the model to focus on just two "principle components" (note that it still incorporates the other variables in the model, but these variables have been "simplified" or "compressed")

With a PCA model, similar results will cluster together, and clusterings can help with classification. With our student example, we would see "A" students cluster together, "B" students to cluster together, etc. While this example might seem like a simple explanation, PCA analysis can be very useful when large amounts of data are available, but you don't know which variables are most important in your classification.

For our purposes, we should be able to use PCA analysis by taking large ammounts of data we collect on tweets (their features, their average word vector, etc.) and see a clustering of complaints and non-complaints, without worrying about which features or variables of a tweet are best for classification.

## Useful Resources:
Quick Explanation of PCA: https://www.youtube.com/watch?v=HMOI_lkzW08
https://www.youtube.com/watch?v=_UVHneBUBW0

# The XGBoost Sub-Model

## Intro to XGBoost
XGBoost is a tool that uses gradient-boosted trees for classification. A basic understanding of decision trees and random forests is crucial to understanding gradient-boosting as well as the XGBoost package. If any of these terms seem unfamiliar, I'd recommend that you look at the links I've posted below.

##### Decision Trees:
Decision trees are models that use a series of binary (yes/no) questions to classify something based on the answers to those binary questions. A decision tree can be randomly generated to model the data. This means that you could have an infinite number of trees that model the same data.

Simple Decision Tree Example:
Is it snowing? If so, it is a good day to ski. If not, it is not a good day to ski.

##### Random Forests:
A single decision tree is not very effective for modeling--singe decision tree models have been known to overfit the data, and therefore innaccurately classify data. However, it should be noted that decision trees are more accurate at classification than a random guess, and this can make them very useful.

Random forests are models that take advantaged of the fact that a decision tree can still provide a better estimate than a random guess, despite a single tree's inaccuracy. Random forests use hundreds of decision trees (each tree randomly generated to model the same data of interest), and classify data based on the consensus of the outputs these trees.

##### Gradient Boosting:
Gradient Boosting builds on the foundation of random forests by optimizing the model. A gradient boosted random forest systematically analyzes the individual trees of a forest that perform poorly, and focuses on cleaning and improving those trees. With many iterations, a boosted model should improve significantly as it gets trained. XGBoost, the package we use for gradient boosting, is simply one popular tool of many for boosted trees.

## Useful Resources:
When I was first learning about XGBoost, I found these youtube videos to be helpful in explaining the very basic concepts behind boosted trees. Hopefully, they can help you too if you're struggling to follow along.

Decision Trees: https://www.youtube.com/watch?v=7VeUPuFGJHk
Random Forests: https://www.youtube.com/watch?v=J4Wdy0Wc_xQ
More Random Forests: https://www.youtube.com/watch?v=nyxTdL_4Q-Q
Cross Validation: https://www.youtube.com/watch?v=TIgfjmp-4BA
Boostrap Aggregating (Bagging): https://www.youtube.com/watch?v=2Mg8QD0F1dQ
MIT lecture on Boosting: https://www.youtube.com/watch?v=UHBmv7qCey4