### 2019-03-27 Code/dallin_code_annotations/02_create_clustering_function.Rmd
Made a folder for annotating Adriel's Code named dallin_code_annotations. I am hoping that doing so will help me better understand the code/model created by Adriel.

---
title: "Create a Clustering Function"
output: github_document
---

# Libraries
```{r}
library(fastrtext)
library(tidyverse)
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
library(parallel)
library(pbapply)
library(stringr)
```

# Overview
In this document, we will be creating several functions that will aid in building our classification model. These functions will be referenced in future scripts to help build our model.

### Average Word Vector Function
This function will take the word vectors for each word in a tweet, and will average the vectors to create a single vector representing the tweet as a whole. The word vectors will be generated by a fasttext unsupervised model. This will effectively become a mathmatical summary of each tweet, and will be a valuable input to our model.
```{r}
# This creates the average word vector to represent each tweet as an embedding vector
avg_word_vec <- function(text, model) {
  # tokenize the text, return a list of tokens
  vec_of_tokens <- strsplit(text, " ")[[1]]
  # apply get_word_vectors to each token in the vector, return matrix of word vecs
  matrix_of_vecs <- get_word_vectors(model, vec_of_tokens)
  # convert matrix to a tibble for easy reduction
  tbl_of_vecs <- tbl_df(matrix_of_vecs) 
  # summarise the word vector tibble into one average tibble
  summarise_all(tbl_of_vecs, funs(mean), na.rm = TRUE)
}
```

### Code I don't understand
```{r}
# I imagine that this somehow takes a look at the nearest neigbors of a single word, or a tweet vector?
 nn_filter <- function(model, token, filter_str, n = 1000) {
  out <- get_nn(model, token, n)
  if(!is.na(filter_str)) {
    tibble(word = names(out), value = out) %>% 
      filter(!grepl(paste0(".*", filter_str, ".*"), word, ignore.case = T))
  } else {
    tibble(word = names(out), value = out)
  }
}

model_eval <- function(token, ..., filter_str = NA) {
  models <- list(...)
  names(models) <- GetDotObjectNames(...)
  cl <- detectCores()
  pblapply(models, nn_filter, token, filter_str, cl = cl)
}

GetDotObjectNames <- function(...) {
  strsplit(
    gsub("[\\(\\)]", 
         "", 
         regmatches(
           deparse(
             substitute(
               list(...))), 
           gregexpr("\\(.*?\\)", 
                    deparse(
                      substitute(
                        list(...)))))[[1]]), 
    split = ", ")[[1]]
}

model_eval_plot <- function(eval_list, n = 20, token, filter_str) {
  out <- lapply(names(eval_list), function(model) {
    eval_list[[model]] %>% 
      mutate(model = model)
  }) %>% 
    bind_rows() %>% 
    group_by(model) %>% 
    top_n(n, value) %>% 
    ungroup() %>% 
    arrange(model, value) %>% 
    mutate(order = row_number())

  out %>% 
    ggplot(aes(order, value, fill = model)) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    facet_wrap(~model, scales = "free") +
    labs(y = "Euclidean Dist",
         x = NULL) +
    ggtitle(paste("Top", n, "Closest Words"), subtitle = paste0("In Euclidean Distance from '", token, "'", ifelse(missing(filter_str) || is.na(filter_str), "", paste0(" with '",filter_str,"' filtered out")))) +
    scale_x_continuous(
      breaks = out$order,
      labels = out$word,
      expand = c(0,0)
    ) +
    coord_flip()
}

model_eval_plot_wrap <- function(token, ..., filter_str = NA, n = 20) {
  model_eval(token, ..., filter_str = filter_str) %>% 
    model_eval_plot(n, token, filter_str)
}

# function factory to compute total within-cluster sum of square, given a list column number
wss_func <- function(data, col_num) {
  function(k) {
    kmeans(unnest(data[,col_num]), k, nstart = 10 )$tot.withinss
  }
}

kmeans_assign <- function(data, kseq, colseq, nstart = 15, iter.max = 30) {
  df_length <- nrow(data)
  data <- filter(data, tweet_text != "")
  len_diff <- df_length - nrow(data)
  if(len_diff > 0) {
    message(paste0("Filtered ", len_diff, " num rows of blank strings"))
  }
  data_colnames <- colnames(data)
  for(i in kseq) {
    for(j in colseq) {
      varname <- paste0("k", i, "_", data_colnames[j])
      data <- mutate(data, !!varname := kmeans(unnest(data[,j]), i, nstart = nstart, iter.max = iter.max)$cluster)
    }
  }
  return(data)
}

sample_num_levels <- function(df, sample_col_num, n = 20) {
  for(i in sample_col_num) {
    col_levels <- max(df[,i])
    for(j in 1:col_levels) {
      out <- 
        df %>% 
        .[which(.[,i] == j),] %>% 
        dplyr::select(tweet_text) %>% 
        sample_n(n, replace = T)
      cat(paste("\n\nSampling", n, "observations from cluster", j, "in column ", colnames(df)[i],"\n\n"))
      print(out)
    }
  }
}
```