### 2019-03-27 Code/dallin_code_annotations/04_test_rds_connection.Rmd

---
title: "Test RDS Connection"
output: github_document
---
# Overview
In this document, I will be annotating the code written by Adriel in the script "TestRDSConnection.R."

```{r}
library(RPostgreSQL)
library(dbplyr)
library(tidyverse)

# Connect to RDS ----------------------------------------------------------

# Load drivers to communicate with RDS
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "MarketingAnalyticsClassData",
                 host = "byumarketinganalytics.chaby6kshgs3.us-east-2.rds.amazonaws.com", port = 5555,
                 user = "adrielc", password = "byuanalytics58") # This will prompt user

# List all tables available within the database
dbListTables(con)

# Create tbl object to reference the table serverside when pulling data
Cockroaches <- tbl(con, "Cockroaches")

# Pull Data Sample --------------------------------------------------------

# Get Tweet text for the first 1,500,000 tweets in the DB. Remove retweets (there's a lot)
sample_tweets <- 
  Cockroaches %>% 
  filter(substr(tweet_text, 1, 4) != "RT @") %>% 
  select(tweet_text) %>% 
  head(1500000) %>% 
  mutate(tweet_text = tolower(tweet_text)) %>%
  collect() # all the processing is done server side before being pulled in by collect()

# Once tweets are pulled in, write the lines to a file.
sample_tweets %>%
  .$tweet_text %>% 
  write_lines("Output/clean_data/tweet_sample_1hM_noRT.txt")

DBI::dbDisconnect(con)

```