### 2019-03-27 Code/dallin_annotations/04_test_rds_connection.Rmd
Wrote more on the test_rds_connection.

---
title: "Test RDS Connection"
output: github_document
---
# Overview
In this document, I will be annotating the code written by Adriel in the script "TestRDSConnection.R."

##### What I understand:
RDS stands for R Data Server. This script helps us log into the RDS and extract the twitter data.

# Libraries
```{r}
library(RPostgreSQL)
library(dbplyr)
library(tidyverse)
```

# Connect to RDS
This code helps us connect to the server and lost the data tables held inside. Our database is apparently a Postgre database
```{r}
# Load drivers to communicate with RDS
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "MarketingAnalyticsClassData",
                 host = "byumarketinganalytics.chaby6kshgs3.us-east-2.rds.amazonaws.com", port = 5555,
                 user = "adrielc", password = "byuanalytics58") # This will prompt user

# List all tables available within the database
dbListTables(con)
```

### Extract the Data
With the code chunk above, we were able to see the data tables listed in the server. The dataframe we are interested in is called "Cockroaches." We will find that table and save our connection to that dataframe. This will make it easy for us to wrangle the data on the server before it arrives on our laptop. This will be advantageous because the server will likely have more computing power than our laptops.
```{r}
# Create tbl object to reference the table serverside when pulling data
Cockroaches <- tbl(con, "Cockroaches")
```

### Pull Data Sample
Now we will use our connection to wrangle and save the data to our computer.
```{r}
# Get Tweet text for the first 1,500,000 tweets in the DB. Remove retweets (there's a lot)
sample_tweets <- Cockroaches %>% 
  filter(substr(tweet_text, 1, 4) != "RT @") %>% 
  select(tweet_text) %>% 
  head(1500000) %>% 
  mutate(tweet_text = tolower(tweet_text)) %>%
  collect() # all the processing is done server side before being pulled in by collect()
```

### Write Tweets to a .txt File
Now that we have wrangled the data and saved it as an object in our environment, we want to use the write_lines() function to save this data file as a .txt file.
```{r}
# Pull Data Sample --------------------------------------------------------
# Once tweets are pulled in, write the lines to a file.
sample_tweets %>%
  .$tweet_text %>% 
  write_lines("Output/clean_data/tweet_sample_1hM_noRT.txt")
```

# Disconnect
It is always important do sever your connection to a database after you use it. 
```{r}
DBI::dbDisconnect(con)
```