### 2019-03-27 Code/dallin_annotations/README.Rmd

---
title: "README"
output: github_document
---
# Dallin Annotations
In this folder, I will keep all of my .rmd documents. Some of them will be created by me, but most will take Adriel's code in other folders and annotate it. I hope that by doing so, I will better understand his code, and be able to contribute more to writing the paper on our reasearch. Additionally, I hope these annotations help future RA's understand the code as well.

The numbering of the files is an attempt to be consistent with the tidyverse style guide section 1.1 (found at https://style.tidyverse.org/). As I work more with the code, I will do my best to ensure that the numbering reflects a logical flow for the code.

If the documents only contain a single, massive block of code, this means that I haven't gotten around to annotating that script.