### 2019-05-08 Code/dallin_annotations/11_build_multi_classifier.Rmd
Built and analyzed a classifier for multiple complaint categories. This uses the same foundation as the "frankenstein model" in file 07_build_classifier.Rmd

---
title: "Building a Frankenstien Classification Model for Multiple Complaint Labels"
output: github_document
---

# Overview
In this document, I will try to make a classification model similar to the one in 07_build_classifier.Rmd. The difference here will be that I will attempt to build a classification model meant to predict based on a set of many complaint categories.

# Libraries and Scripts
```{r echo = FALSE}
library(tidyverse)
library(fastrtext)
library(magrittr)
library(plotly)
require(randomForest)
library(Matrix)
library(xgboost)
library(caret)
library(car)

set.seed(42)

source("05_define_clustering_functions.R")
source("06_define_prediction_functions.R")
```

## Load the Training Dataset
This dataset was created in file 07_build_classifier.Rmd
```{r}
load(here::here("Output", "train_data.dat"))
```

Modify the training dataset to make it ready for a multi classification
```{r}
train <- train %>% 
  mutate(complaint_category_name = complaint_category) %>% 
  mutate(complaint_category = case_when(complaint_category_name == "non_complaint"         ~ 1,
                                        complaint_category_name == "delay_delay"           ~ 2,
                                        complaint_category_name == "event_specific"        ~ 3,
                                        complaint_category_name == "event_vague"           ~ 4,
                                        complaint_category_name == "human_service_rude"    ~ 5,
                                        complaint_category_name == "baggage"               ~ 6,
                                        complaint_category_name == "app_failure"           ~ 7,
                                        complaint_category_name == "delay_cancel"          ~ 8,
                                        complaint_category_name == "human_service_hold"    ~ 9,
                                        complaint_category_name == "delay_tarmac"          ~ 10,
                                        complaint_category_name == "wishlist"              ~ 11,
                                        complaint_category_name == "check_failure"         ~ 12))
```

I will make sure that the mutated data matches the original data.
```{r}
train %>% 
  count(complaint_category_name, complaint_category, sort = TRUE) %>% 
  mutate(complaint_category_name = reorder(complaint_category_name, n))
```
We see that every datapoint has been matched, so we're good!

## Partition the Dataset
```{r}
trainIndex <- createDataPartition(train$complaint_category, p = .8, 
                                  list = FALSE, 
                                  times = 1)

test_data <- train[-trainIndex,]
test_labels <- test_data %>% select(complaint_category) %>% transmute(label = complaint_category)
test_data_vec <- test_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)

train_data <- train[trainIndex,]
train_labels <- train_data %>% select(complaint_category) %>% transmute(label = complaint_category)
train_data_vec <- train_data %>%  select(tweetFeatures, avg_vec_50, avg_vec_50_PCA) %>% unnest %>% mutate_all(as.numeric)
```

## Create Dense Matricies for XGBoost
```{r}
dtest1 <- xgb.DMatrix(sparse.model.matrix(~.-1, test_data_vec))
```

## Set the Parameters of the Model and Begin Training
```{r error = TRUE}
bst <- xgboost(data = sparse.model.matrix(~.-1, train_data_vec), 
               label = as.factor(train_labels$label),
               max.depth = 4, 
               eta = 0.2, 
               nthread = 2, 
               nrounds = 100, 
               num_class = 13,
               objective = "multi:softmax",
               eval_metric = evalerror)
```
The train-error does not improve with iterations, which is concerning to me.

`# Get Predictions on the Model
```{r error = TRUE}
test_data$prediction <- predict(bst, dtest1)

predict(bst, dtest1)
```

!!!!! I messed up on the predictions and will need to come back to this to see how to fix it.

```{r error = TRUE}
test_data %>% 
  ggplot(aes(x = as.factor(complaint_category), fill = as.factor(prediction))) +
  geom_bar() +
  coord_flip()
```
We see that the model isn't especially accurate, but also that for many test labels, the majority of the predicitions (that aren't non complaints) are correct (i.e. 1 matches 1, 2 matches 2, 4 matches 4, etc.). The model could definitely be improved, but it performs better than rolling the dice.

## Accuracy
```{r}
test_data <- test_data %>% 
  mutate(correct = complaint_category == prediction)

test_data %>% 
  summarise(mean(correct))
```
The model is 76% accurate, which is suprisingly good! It definitely has room for improvement though. I would guess that the model's ability to correctly estimate the majority of non complaints has helped the model perform well.

How does the model perform when we focus solely on complaints though?

Visualization of complaint categories:
```{r}
test_data %>% 
  filter(complaint_category != 1) %>% 
  ggplot(aes(x = as.factor(complaint_category), fill = as.factor(prediction))) +
  geom_bar() +
  coord_flip()
```

Accuracy of non-complaint predictions:
```{r}
test_data %>% 
  filter(complaint_category != 1) %>% 
  summarise(mean(correct))
```
It appears that when predicting non-complaints, the model is only accurate about 25% of the time. This is very poor, but better than rolling the dice.

```{r error = TRUE}
confusionMatrix(factor(test_data$complaint_category), factor(test_data$prediction))
```