### 2019-05-08 Code/dallin_annotations/10_build_fasttext_multi_classifier.Rmd
Built a fastText classifier for multiple complaint categories (similar to the classifier built in file 04). This model isn't very accurate, but it's a start.

---
title: "Building a fastText Classifier for Multiple Labels (Supervised)"
output: github_document
---
# Overview
In this document, I create a fastText classification model that predicts complaints by the various complaint categories we created in our test/train dataset. This model is very simple, and uses fastText exclusively for classification.

## Load Packages
```{r}
library(fastrtext)
library(tidyverse)

set.seed(42)
```

Set the paths to the model, testing, and training data. Once the data and models are created, this is where they will be saved.
```{r}
model_path <- here::here("Output", "models", "fasttext_multi_class_model.dat")
train_data_path <- here::here("Output", "clean_data", "rain_multi_data_1.txt")
test_data_path <- here::here("Output", "clean_data", "test_multi_data_1.txt")
```

## Read in the Data
```{r}
tweet_data <- read_csv(here::here("Temporary", "Marketing Research Labeled Tweets_ - tweet_sample_5k_Ky-Ch-Ad.csv")) %>%
 filter(nchar(tweet_text) > 5)
```

## Partition Data for Testing and Training
```{r}
test_train_data <- tweet_data %>%
  mutate(input_lines = paste(complaint_category, tweet_text)) %>%
  group_by(complaint_category) %>% # get a representative sample of both complaints and non complaints
  mutate(test_train = c("test", "train")[rbinom(n(), 1, 0.8) + 1]) %>% # assign test train splits
  ungroup()
```

## Create .txt Files
As input, the fastText model needs to take the testing and training data as .txt files.
```{r}
# write train data
train_data_lines <- test_train_data %>%
  filter(test_train == "train") %>%
  pull(input_lines) %>%
  paste0("__label__", .)

write_lines(train_data_lines, train_data_path)

# write test data
test_data_lines <- test_train_data %>%
  filter(test_train == "test") %>%
  pull(input_lines) %>%
  paste0("__label__", .)

write_lines(test_data_lines, test_data_path)

# create a character vector containing the tweets to test without their labels
test_labels_without_prefix <-
  test_train_data %>%
  filter(test_train == "test") %>%
  pull(complaint_category)
```

## Train the fastText Classification Model (Supervised, Multiple Categories)
```{r}
# train model
execute(commands = c("supervised", "-input", train_data_path, "-output", model_path, "-dim", 20, "-lr", 1, "-epoch", 20, "-wordNgrams", 2, "-verbose", 1))

# load model
model <- load_model(model_path)
```

## Review Model
```{r}
predictions <- predict(model, sentences = test_data_lines, simplify = TRUE, unlock_empty_predictions = TRUE)

head(predictions, 5)
```

Below we compare the length of the list of predictions to the length of the testing dataset. Ideally, these should have the same length. This will help us confirm that our model worked as expected.
```{r}
length(predictions)
length(test_labels_without_prefix)
length(test_data_lines)
```
Similar to the previous fastText model, it appears the model would not make predictions on a few of the tweets in the testing dataset. This is likely due to the size of our training dataset as well as the variance in vocabulary for tweets.

Below, we merge the predictions with our testing dataset, preparing the data for visualizations.
```{r}
new_predictions <- attributes(predictions)$names

tested <- test_train_data %>% 
  filter(test_train == "test") %>% 
  mutate(prediction = predict(model, tweet_text, unlock_empty_predictions = TRUE)) %>% 
  unnest(prediction) %>% 
  mutate(new_predictions = new_predictions,
         correct = case_when(complaint_category == new_predictions ~ TRUE,
                             complaint_category != new_predictions ~ FALSE))
```

## Visualize Predictions
```{r}
tested %>% 
  ggplot(aes(x = complaint_category, fill = new_predictions)) +
  geom_bar() +
  coord_flip()
```

## Accuracy
```{r}
tested %>% 
  summarise(accuracy = mean(correct))
```
This model does not seem very accurate at all, but it is still better than rolling the dice.