### 2019-10-17 Code/postratification_code.Rmd
Begginings of a postratification model for our MRP model.

---
title: "Postratification Code"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Postratification code

We are looking into an MRP model for Counting Cockroaches and we want to learn what postratification is and if we can do it in our model.

The documents we study are currently together on a google doc

```{r}
library(tidyverse)

vote_yes <- c(rep(0, 25*(1-0.76)), 
              rep(1, 25*0.76),
              rep(0, 50*(1-0.3)),
              rep(1, 50*0.3))
vote_yes

```