### 2020-03-20 Code/mrp/Case Study .Rmd
This is a working hierarchal multinomial logit model. I am working on making it more interpretable. This comes from the ART forum 2017.

---
title: "MRP Case Study"
author: "McKenna Weech, Adam Hettinger, and Hannah Thompson"
date: "1/17/2020"
output: github_document
---

Firms employ a variety of passive listening platforms (e.g., customer comment cards, help lines, social media) that elicit a variety of complaints. Since these complaints are self-selected, it isn’t obvious if complaints are widespread or genuinely indicative of systemic service failures. In other words, how can can a firm know when complaints can be ignored and when they need to addressed? This issue is related to a larger class of problems where the goal is to make inference about the size of a population from a non-random sample (e.g., the German tank problem). In this paper, we develop a model that uses a variety of observed customer complaints on Twitter to make inference about the severity of service failures.

## Describe the model conceptually

It has been said that if you observe a cockroach on your floor, there are likely thousands inside the walls of your home. Following this analogy, our hope is to count the cockroaches we can see and calculate how many are hiding in the walls.

For the first portion of this case, we will be making a model around simulated data. Each observation will represent a single tweet, and we will be scoring them for satisfaction. We will use a multiple hierarchal linear model to predict satisfaction from the other known variables. Once we have the model predicting well on our simulated data, we will fit it to our real data, and from there use our real data satisfaction scores to measure service or product failure severity.

Along with making a model, we will also be using poststratification to correct the weights in our simulated sample and real data to more correctly represent the population. This will be explained in greater detail as we postratify in this first iteration of the model, so we can explain and see the benefit of a full multiple regression with postratification model (MRP).

To begin, we will start by generating our simulated data. It will run through an initial simple linear model and postratification.

# Iteration 1: Simulation Data and Simple Model

This initial block of code, we will just be loading in the needed packages and settings to make our sample and model code run.

```{r message = FALSE}

library(tidyverse) #Load packages
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)
theme_set(bayesplot::theme_default())
library(dplyr)
library(tidyr)
library(tidybayes)
options(mc.cores = parallel::detectCores()) # Set Stan to use all availible cores
rstan_options(auto_write = TRUE) # Don't recompile Stan code that hasn't changed
```

For this first iteration of modeling, we will be generating our own data. We will generate a population, and sample from it in order to build an MRP model.

We will be modeling our populations `service_failure`. This will eventually translate into measuring service severity when we apply our model to our real data. We will be looking at common characteristics like `gender`, `ethnicity`, `income`, `age`, and `state`.

The following code will make a function `simulate_mrp_data` that will help us make our simulated data, and sample from it in such a way that we can use and see the value in postratification.

```{r}
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
  }
  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)
  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }
  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }
  #male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]
  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))
  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r}
mrp_sim <- simulate_mrp_data(n=1200)
sample <- mrp_sim[["sample"]]
poststrat <- mrp_sim[["poststrat"]]
true_popn <- mrp_sim[["true_popn"]]
head(sample)
```

```{r}
stan_data <- list( 
  N = nrow(sample), 
  service_failure = sample$service_failure,
  male = sample$male, 
  age = sample$age,
  eth = sample$eth, 
  income = sample$income, 
  state = sample$state)
```

We now have need for a logistic regresson model in order to predict a binary outcome for 'service_failure' so we will make a hierarchal logit model

# Hierarchal Logit Model

```{}

data {
  int<lower=2> C; // # of alternatives (choices) in each scenario
  int<lower=1> K; // # of covariates of alternatives
  int<lower=1> R; // # of respondents
  int<lower=1> S; // # of scenarios per respondent
  int<lower=0> G; // # of respondent covariates 
  int<lower=1,upper=C> Y[R, S]; // observed choices
  matrix[C, K] X[R, S]; // matrix of attributes for each obs
  matrix[G, R] Z; // vector of covariates for each respondent
}

parameters {
  matrix[K, R] Beta;
  matrix[K, G] Theta;
  corr_matrix[K] Omega;
  vector<lower=0>[K] tau;
}

transformed parameters {
  cov_matrix[K] Sigma = quad_form_diag(Omega, tau);
}

model {
  //priors
  to_vector(Theta) ~ normal(0, 10);
  tau ~ cauchy(0, 2.5); 
  Omega ~ lkj_corr(2);
  //likelihood
  for (r in 1:R) {
    Beta[,r] ~ multi_normal(Theta*Z[,r], Sigma);	
    for (s in 1:S)
      Y[r,s] ~ categorical_logit(X[r,s]*Beta[,r]);
  }
}

```

```{r}
# ========== Load libraries and set options ==========
library(rstan)
library(MASS)
library(shinystan)

# writes a compiled Stan program to the disk to avoid recompiling
rstan_options(auto_write=TRUE) 
# allows Stan chains to run in parallel on multiprocessor machines
options(mc.cores = parallel::detectCores())

# ========== Test Stan with synthetic data ============

# function to generate mnl data
generate_hmnl_data <- function(R=100, S=30, C=3, 
                               Theta=matrix(rep(1, 8), nrow=2), 
                               Sigma=diag(0.1, 4)){
  K <- ncol(Theta)
  G <- nrow(Theta)
  Y <- array(dim=c(R, S))
  X <- array(rnorm(R*S*C*K), dim=c(R, S, C, K)) # normal covariates
  Z <- array(dim=c(G, R))
  Z[1,] <- 1  # intercept
  if (G > 1) {
    Z[2:G,] <- rnorm(R*(G-1)) # normal covariates
  }
  Beta <- array(dim=c(K, R))
  for (r in 1:R) {
    Beta[,r] <- mvrnorm(n=1, mu=Z[,r]%*%Theta, Sigma=Sigma)
    for (s in 1:S)
      Y[r,s] <- sample(x=C, size=1, prob=exp(X[r,s,,]%*%Beta[,r])) # logit formula
   }
  list(R=R, S=S, C=C, K=K, G=G, Y=Y, X=X, Z=Z, 
       beta.true=beta, Theta.true=Theta, Sigma.true=Sigma)
}

d1 <- generate_hmnl_data()
str(d1)

test.stan <- stan(file="hmnl.stan", data=d1, iter=1000, chains=4)

plot(test.stan, plotfun="trace", pars=("Theta"))

plot(test.stan, plotfun="trace", pars=c("tau"))

plot(test.stan, plotfun="trace", pars=("Omega"))

plot(test.stan, pars=c("Theta", "tau", "Omega"))

plot(test.stan, pars=c("Theta", "Sigma"))

summary(test.stan, pars=c("Theta"))$summary

summary(test.stan, pars=c("tau"))$summary

summary(test.stan, pars=c("Omega"))$summary
```

```{r}
# ========= Read in Chocolate Data and Prep for Stan ==========
rm(list=ls()) # tidy up
choc.df <- read.csv("cbc_chocolate.csv")

# Coding the chocolate data (this ought to be a function)
choc.contrasts <- list(Brand = "contr.sum", Type = "contr.sum")
choc.coded <- model.matrix(~ Brand + Type, data = choc.df, contrasts = choc.contrasts)
choc.coded <- choc.coded[,2:ncol(choc.coded)] # remove intercept
# Fix the bad labels from contr.sum
choc.names <- c("BrandDove", "BrandGhirardelli", "BrandGodiva", "BrandHersheys", 
                "TypeDark", "TypeDarkNuts", "TypeMilk", "TypeMilkNuts")
colnames(choc.coded) <- choc.names
choc.df <- cbind(choc.df, choc.coded)

head(choc.df)

# Munge into Stan list
R <- length(unique(choc.df$Ind))
S <- length(unique(choc.df$Trial))
C <- max(choc.df$Alt)
K <- 9
Y <- array(dim=c(R, S))
X <- array(rnorm(R*S*C*K), dim=c(R, S, C, K)) 
Z <- array(1, dim=c(1, R)) # intercept only
for (r in 1:R) { # respondents
  for (s in 1:S){ # choice scenarios
    scenario <- choc.df[choc.df$Ind==unique(choc.df$Ind)[r] & 
                        choc.df$Trial==unique(choc.df$Trial)[s], ]
    X[r,s,,] <- data.matrix(scenario[,c(7, 9:16)]) # price and coded brand and type
    Y[r,s] <- scenario$Alt[as.logical(scenario$Chosen)]
  }
}

choc.standata <- list(C=C, K=K, R=R, S=S, G=1, Y=Y, X=X, Z=Z)
str(choc.standata)

rm(Y, X, Z, R, S, r, s, C, K, choc.contrasts, scenario, choc.coded)

```

```{r}

# ========== Run Stan and Check Convergence =========
choc.stan <- stan(file="hmnl.stan", data=choc.standata)

plot(choc.stan, plotfun="trace", pars=("Theta"))

plot(choc.stan, plotfun="trace", pars=c("tau"))

plot(choc.stan, plotfun="trace", pars=c("Omega[1,2]"))

plot(choc.stan, plotfun="trace", pars=paste("Beta[", 1:9, ",1]", sep="")) # resp 1

summary(choc.stan)$summary[,c("Rhat", "n_eff")]

# Convergence checking gets tedious with a lot of parameters, so automate
check_fit <- function(fit) {
  summ <- summary(fit)$summary
  range_rhat <- range(summ[ , 'Rhat'])
  rhat_ok <- 0.99 <= range_rhat[1] && range_rhat[2] <= 1.1
  range_neff <- range(summ[ , 'n_eff'])
  neff_ok <- range_neff[1] >= 400
  sp <- rstan::get_sampler_params(fit, inc_warmup=FALSE)
  max_divergent <- max(sapply(sp, function(p){ sum(p[ , 'divergent__']) }))
  no_divergent <- max_divergent == 0

  list(ok = rhat_ok && neff_ok && no_divergent,
       range_rhat = range_rhat,
       range_neff = range_neff,
       max_divergent = max_divergent)
}

check_fit(choc.stan)
```

```{r}
# ========== Summarize Posterior ==========
choc.names
plot(choc.stan, pars=c("Theta", "tau"))

plot(choc.stan, pars=c("Omega"))

plot(choc.stan, pars=paste("Beta[", 1:9, ",1]", sep="")) + ggtitle("Respondent 1: Likes Milk Chocolate")

plot(choc.stan, pars=paste("Beta[", 1:9, ",2]", sep="")) + ggtitle("Respondent 2: Likes Dark Chocolate")

launch_shinystan(choc.stan)

# ========== Simulate Shares from Model =========
# function for computing shares from beta.draws for an hmnl model
shares.hmnl.post <- function(beta.draws, X) # X is attribute matrix for scenario
{
  R <- dim(beta.draws)[3]  # respondents
  D <- dim(beta.draws)[1]  # draws
  shares <- array(NA, dim=c(nrow(X), R, D))
  for (d in 1:D) {
    for (r in 1:R) {
      beta <- beta.draws[d,,r] 
      V <- exp(X %*% beta)  
      shares[,r,d] <- V/sum(V)
    }
  }
  shares
}

choc.standata$X[1,1,,] # note option 2 is dark
shares <- shares.hmnl.post(extract(choc.stan, pars=c("Beta"))$Beta, 
                          choc.standata$X[1,1,,])
str(shares)

apply(shares, 1, quantile, probs=c(0.5, 0.025, 0.975))
apply(shares, 1:2, quantile, probs=c(0.5, 0.025, 0.975))

```