### 2020-03-20 Code/mrp/brms code.Rmd

---
title: "Why BRMS Doesn't Work"
author: "McKenna Weech"
date: "2/27/2020"
output: word_document
---

```{r}
# Load packages
library(brms)
```

```{r}
#Let's grab some data 
sample <- mrp_sim[["sample"]]

#And let try and get data from brms.
make_standata(formula =
                cat_pref ~ factor(male) + factor(male) * factor(age) + # gender factor and a gender ~ age interaction factor
                (1 | state) + (1 | age) + (1 | eth) + (1 | income), # random intercept with fixed means
              family = binomial(link = "logit"),
              data = sample)
```

We can also see the stan code that it is writing. 
```{r}
make_stancode(formula =
                cat_pref ~ factor(male) + factor(male) * factor(age) + # gender factor and a gender ~ age interaction factor
                (1 | state) + (1 | age) + (1 | eth) + (1 | income), # random intercept with fixed means
              family = binomial(link = "logit"),
              data = sample )
```