### 2020-05-06 Code/mrp/Case Study (McKenna) .Rmd
Inserted downes mrp model as well as the stan user guide mrp model. Edited for case study format.

---
title: "Case Study (McKenna)"
author: "McKenna Weech"
date: "3/10/2020"
output: html_document
---

## Intro 
Multi-level regression and post-stratification helps to infer treatment size for a whole population from a non-representative sample. Using population knowlege sample are classified into cells baised off of choosen classification variables. Heirarchical bayesian models are used to infer(?) treatment within cells and population knowlege is used to weight each cell according to their prevelance within the total population.

## Data/Categories/Cells
Categorical classification variables are choosen and cell are assigned baised off of category classification. For example, if one variable is gender consisting of two categories (male and other) and one varibale is age consisting of two categories (0-50 and 50-100) then there would be four cells: male 0-50, male 50-100, other 0-50, and other 50-100.

In this case study there are four classification variables choosen: age, income, ethnicity, gender, and state. Each vairbale is tranforemed into categorical variables with age having seven categories, income having three, ethnicity having three, gender having two, and state have 50 making the total number of cells 6,300  (7 x 3 x 3 x 2 x 50 = 6300).

Data is generated using a function from...(find referance)

## Model

```{r message = FALSE, echo = FALSE}
library(tidyverse) #Load packages
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)
theme_set(bayesplot::theme_default())
library(dplyr)
library(tidyr)
library(tidybayes)
library(here)

options(mc.cores = parallel::detectCores()) # Set Stan to use all availible cores
rstan_options(auto_write = TRUE) # Don't recompile Stan code that hasn't changed
```

```{r, eval = FALSE, echo = FALSE}
# Data generating function for referance. 
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
  }

  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)

  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }

  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }

  #male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]

  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))

  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r, message = FALSE, echo = FALSE}

# Import data. 
sample <- read_csv(here::here("code", "mrp", "data", "sample.csv"))
proststrat <- read_csv(here::here("code", "mrp", "data", "poststrat.csv"))
true_popn <- read_csv(here::here("code", "mrp", "data", "true_popn.csv"))

```

This case study will use generated data. The data is broken up into three data tables: sample, postratification, and population. The postratification table shows the population density of each cell and is used to weight the outcome results from the sample. Because the data is generated we also have a generated population that we can use to check the accuracy of the model.

First we will build a model to predict the our outcome for each cell. Then we will use the postratification table to correctly weight each cell to predict outcome for the whole population.

We will use bayesian hierchical modeling to take advantage of partical pooling to account for empty cells.

#### McKenna's logit model  
```{r}
data_list <- list(
  D <- 5,                                                    # Number of variables. 
  N <- 1200,                                                 # Number of observations.
  L <- 6300,                                                 # Number of groups. 
  y <- sample$service_failure,                               # Outcome variables. 
  ll <- sample(6300, 1200, replace = TRUE),                  # Group assignment. 
  x <- matrix(sample %>% 
                select(male, age, eth, income, state))       # Matrix of predictors. 
)

D <- 5
N <- 1200
L <- 6300
y <- sample$service_failure
ll <- sample(6300, 1200, replace = TRUE)

```

```{r, eval = FALSE}
fit <- stan(
  file = here::here("Code", "mrp", "logit_hier.stan"),
  data = "data_list",
  seed = 42
)
```
#### Downes logit model 
```{r}
# Observations. 
outcome <- sample$service_failure
male <- sample$male
age <- sample$age 
eth <- sample$eth 
income <- sample$income
state <- sample$state

# Number of categories. 
n <- length(outcome)
n_male <- max(male)
n_age <- max(age)
n_eth <- max(eth)
n_income <- max(income)
n_state <- max(state)

# Define parameters for Hamiltonian Monte Carlo algorithm
n_iter <- 1000
n_chain <- 4
n_sim <- (n_iter * n_chain)/2

```

```{r, eval = FALSE}
stanfit_final <- stan(file= "downes_copy.stan",
                        data=c("n","n_age","n_male","n_eth","n_income","n_state",
                               "outcome","age","male","eth","income","state"), 
                        iter=n_iter, chains=n_chain)
```
#### Stan user guide logit model 
```{r}
stan_data <- list( 
  age <- sample$age, 
  eth <- sample$eth, 
  income <- sample$income, 
  male <- sample$male, 
  state <- sample$state, 
  Y <- sample$service_failure, 
  N <- length(Y))
```

```{r, eval = FALSE}
stanfit_final <- stan(file= "stan_mrp.stan",
                        data= "stan_data", 
                        iter=n_iter, chains=n_chain)
```

## Postratification

Postratification consists of taking the sum of the estimates of the model for each cell times the number of people in each cell divided by the total number of people

$$\omega=\displaystyle\frac{$$\sum_{j=1}^{J}\beta_{j}*N_{j}$$}{$$\sum_{j=1}^{J}N_{j}$$}$$
Postratification can be performed in stan using the generated quantities block.

```{r, eval = FALSE}
generated quantities {
  real expect_pos = 0;
  int total = 0;
  for (b in 1:7)
    for (c in 1:3)
      for (d in 1:3)
        for (f in 1:2)
          for (g in 1:50) {
        total += P[b, c, d, f, g];
        expect_pos
          += P[b, c, d, f, g]
             * inv_logit(b0 + a_age[b] + a_eth[c] + a_income[d] + a_male[f] + a_state[g]);
      }
  real<lower = 0, upper = 1> phi = expect_pos / total;
}           
```

## Compare to Population 
Because we are using simulated data we can see how our poststratified predictions compare to the true population