### 2020-05-06 Code/MRP-Hannah/Citations.Rmd

---
title: "Citations"
author: "Hannah Thompson"
date: "5/6/2020"
output: html_document
bibliography: bibliography.bibtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

https://timmastny.rbind.io/blog/multilevel-mrp-tidybayes-brms-stan/  
@misc{ Mastny17
       author = "Tim Mastny"
       title = "MRP Using brms and tidybayes",
       year = "2017" }

https://timmastny.rbind.io/blog/poststratification-with-dplyr/ 
@misc{ @Mastny17,
       author = "Tim Mastny",
       title = "Poststratification Primer with dplyr",
       year = "2017" }

https://cran.r-project.org/web/packages/rstanarm/vignettes/mrp.html
@misc{ @Kennedy_Gabry20,
       author = "Lauren Kennedy", "Jonah Gabry",
       title = "MRP with rstanarm",
       year = "2020" }

https://github.com/gelman/mrp-stan
@misc{ @Gelman14,
       author = "Andrew Gelman",
       title = "gelman/mrp-stan",
       year = "2014" }

https://www.r-bloggers.com/gelmans-mrp-in-r-what-is-this-all-about/
@misc{ Gutiérrez17,
       author = "Gutiérrez, Andrés",
       title = "Gelman's MrP in R – What is this all about?",
       year = "2017" }

http://www.stat.columbia.edu/~gelman/research/unpublished/MRT(1).pdf
@misc{ @gelman_lax_phillips_gabry_trangucci_2018,
       author = "Gelman, Andrew and Lax, Jeffrey and Phillips, Justin and Gabry, Jonah and Trangucci, Robert",
       title = "Using Multilevel Regression and Poststratification to Estimate Dynamic Public Opinion",
       year = "2018" }

https://github.com/gelman/mrp-stan/blob/master/models/turnout_simple_post.stan
@misc{ @Gelman14,
       author = "Andrew Gelman",
       title = "gelman/mrp-stan",
       year = "2014" }

http://www.stat.columbia.edu/~gelman/research/unpublished/modelweighting.pdf
@misc{ @si_trangucci_gabry_gelman_2018,
       author = "Si, Yajuan and Trangucci, Rob and Gabry, Jonah and Gelman, Andrew",
       title = "Bayesian hierarchical weighting adjustment and survey inference",
       year = "2018" }

https://cran.r-project.org/web/packages/rstanarm/vignettes/pooling.html
@misc{ @carpenter_gabry_goodrich_2020,
       author = "Carpenter, Bob and Gabry, Jonah and Goodrich, Ben",
       title = "Hierarchical Partial Pooling for Repeated Binary Trials",
       year = "2020" }

https://academic.oup.com/aje/article/187/8/1780/4964985
@article{ @Nobody06,
       author = "Downes, Marnie and Gurrin, Lyle C and English, Dallas R and Pirkis, Jane and Currier, Dianne and Spittal, Matthew J and Carlin, John B",
       title = "Multilevel Regression and Poststratification: A Modeling Approach to Estimating Population Quantities From Highly Selected Survey Samples",
       year = "2018" }

https://github.com/danilofreire/r-scripts/blob/master/stan-logistic-regression.R 
@misc{ @kennedy_gabry_2019,
       author = "Kennedy, Lauren and Gabry, Jonah",
       title = "MRP with rstanarm",
       year = "2019" }

https://mc-stan.org/docs/2_18/stan-users-guide/hierarchical-logistic-regression.html
@misc{ @Freire15,
       author = "Danila Freire",
       title = "Stan Logistic Regression",
       year = "2015" }

https://mc-stan.org/docs/2_18/stan-users-guide/hierarchical-logistic-regression.html

@manual{ @StanGuide,
       author = "Team, Stan Development",
       title = "Stan User's Guide" }