### 2020-09-29 Cases/mrp/Case Study (McKenna) .Rmd
Updating poststratification.

---
title: "Case Study (McKenna)"
author: "McKenna Weech, Hannah Thompson"
date: "3/10/2020"
output: github_document
---

Multilevel regression and post-stratification helps to infer treatment size for a whole population from a non-representative sample. Using population knowledge, samples are classified into cells based off of chosen classification variables. Hierarchical Bayesian models are used to infer treatment within cells and population knowledge is used to weight each cell according to their prevalence within the total population. Multilevel regression and post-stratification (Gelman & Little, 1997) proceeds by fitting a hierarchical regression model to survey data, and then using the population size of each post-stratification cell to construct weighted survey estimates. The accuracy of post-stratification within MRP can be dependent on whether the matrix that is used in post-stratification is an accurate representation of the target population. This combination of multilevel regression and post-stratification has been seen as an effective method when adjusting a sample to be more representative of a population for specific variables.

## Data

Categorical classification variables are chosen and cells are assigned based off of category classification. For example, if one variable is gender consisting of two categories (male and other) and one variable is age consisting of two categories (0-50 and 50-100) then there would be four cells: male 0-50, male 50-100, other 0-50, and other 50-100.

In this case study there are five classification variables: age, income, ethnicity, gender, and state. Each variable is transformed into categorical variables with age having seven categories, income having three, ethnicity having three, gender having two, and state have 50 making the total number of cells 6,300 (7 x 3 x 3 x 2 x 50 = 6300).

Data is generated using a function used in case study (noted in references).

## Model

```{r load-data, message=FALSE}
library(tidyverse) #Load packages
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)
theme_set(bayesplot::theme_default())
library(dplyr)
library(tidyr)
library(tidybayes)
library(here)
```

```{r, eval = FALSE}
# I set this to eval = FALSE just for now because it wasn't knitting so I will have to come back and take a look at this.
options(mc.cores = parallel::detectCores()) # Set Stan to use all available cores.
rstan_options(auto_write = TRUE)            # Don't recompile Stan code that hasn't changed.

# Data generating function for reference. 
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) 
  # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) 
#Adjust the N to be the number observed in each category in each group
  }
# Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)
  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }
  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }
#male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]
  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))
  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r}
# Importing the data. 
sample <- read_csv(here::here("Data", "sample.csv"))
poststrat <- read_csv(here::here( "Data", "poststrat.csv"))
true_popn <- read_csv(here::here( "Data", "true_popn.csv"))

# Add groups to data. 
poststrat_group <- poststrat %>% 
  mutate(group = row_number())

sample_group <- left_join(sample,
                          poststrat_group,
                          by = c(
                            "male" = "male", 
                            "age" = "age", 
                            "eth" = "eth", 
                            "income" = "income",
                            "state" = "state") )
```

This case study will use generated data. The data is broken up into three data tables: sample, post-stratification, and population. The post-stratification table shows the population density of each cell and is used to weight the outcome results from the sample. Because the data is generated we also have a generated population that we can use to check the accuracy of the model.

First we will build a model to generate parameter estimates for each cell. Then we will use the post-stratification table to correctly weight the estimates for the whole population.

We will use Bayesian hierarchical modeling to take advantage of partial pooling to account for underweight cells.

### McKenna's logit model

Our first model is a hierarchical logit model. A logit model is appropriate for our data because of the biniary outcome variable. The hierarchical model is group baised on the group assignments determined by the categorical variables in the data.

```{stan logit-hier, output.var="logit_hier", eval=FALSE}
data {
  int<lower=1> D;                // Number of variables.
  int<lower=0> N;                // Number of observations.
  int<lower=1> L;                // Number of groups.
  int<lower=0,upper=1> y[N];     // Outcome variable. 
  int<lower=1,upper=L> ll[N];    // Group assignment. 
  matrix[N, D] x;                // Matrix of predictors. 
}

parameters {
  real mu[D];
  real<lower=0> sigma[D];
  vector[D] beta[L];
}

model {
  mu ~ normal(0, 100);
  sigma ~ normal(0, 5);
  for (l in 1:L) {
    beta[l] ~ normal(mu, sigma);
  }
  for (n in 1:N) {
    y[n] ~ bernoulli_logit((x[n,] * beta[ll[n]]));
  }
}
```

...THEN YOU CAN FIT.

```{r}
predictors <- sample_group %>% 
  select(male, age, eth, income, state)

data_list <- list(
  D = 5,                                              # Number of variables. 
  N = 1200,                                           # Number of observations.
  L = 6300,                                           # Number of groups. 
  y = sample_group$service_failure,                   # Outcome variables. 
  ll = sample_group$group,                            # Group assignment. 
  x = predictors                                      # Matrix of predictors. 
)
```

```{r, eval = FALSE}
fit <- stan(
  file = here::here("Code", "mrp", "logit_hier.stan"),
  data = data_list,
  seed = 42 
)
```
```{r, eval = FALSE}
# I think we will need to use a for loop like they do in the downes model to construct our predictions into a format that we can then postratify them. I got this for loop from the downes code and I have just begun to adjust it for our model. 
 n_sim <- (n_iter * n_chain)/2   # They set the chains in their code but I don't think we did for our model so we'll have to take a look and see how to convert this 
n_pscell <- length(poststrat$N) # (n_pscell is the number of poststratification cells defined by the model)
n_state <- max(poststrat$STATE)
stanfit_final_ypred <- array(NA, c(n_sim, n_pscell))

for(l in 1:n_pscell){
    stanfit_final_ypred[,l] <- bernoulli_logit((x[n,] * beta[ll[n]])  # This is the part I am confused the most on. Just trying to specify the model in the right way to be able to make the service failure assumptions to then postratify. 
                                        )

stanfit_final_ps <- c(NA, n_sim)

# THEN WE CAN FINALY POSTSTRATIFY....I think 
for(s in 1:n_sim){
    stanfit_final_ps[s] <- sum(poststrat$N*stanfit_final_ypred[s,])/sum(poststrat$N)
  }

# This is just some other code that I am working on to try and look at a different way to post. I think this mainly comes from the original case study from Lauren. 
posterior_prob <- posterior_linpred(fit, transform = TRUE, newdata = poststrat)

poststrat_prob <- posterior_prob %*% poststrat$N / sum(poststrat$N)

model_popn_pref <- c(mean = mean(poststrat_prob), sd = sd(poststrat_prob))

round(model_popn_pref, 3)
```

### Downes logit model

The data section of the Downes model not only specifies the individual predictor values but also the number of categories for each categorical variable. Each category is given a varying coefficient in the parameter section.

[Note to McKenna]
Questions to answer: 
Why cauchy?
Actual model specification? - looks like the model is specified in the transformed parameters section. why?
Why varying coefficients?

```{stan downes, output.var="downes", eval=FALSE}
data {
  int<lower=0> n;                                     // Number of survey participants

  int<lower=0> n_age;                                 // These are the number of categories
  int<lower=0> n_eth;                                 // for each poststratification factor
  int<lower=0> n_income;                              // in the final model
  int<lower=0> n_male;                                   
  int<lower=0> n_state;

  int<lower=0,upper=1> outcome[n];                    // observed outcome data, 1=Yes, 0=No

  int<lower=0,upper=n_age> age[n];                    // These are the variables containing
  int<lower=0,upper=n_eth> eth[n];                    // the observed data for each
  int<lower=0,upper=n_income> income[n];              // poststratification factor in the 
  int<lower=0,upper=n_male> male[n];                  // final model
  int<lower=0,upper=n_state> state[n];      
 }
parameters {                                          // Model parameters to be estimated:
  real b0;                                            // "intercept"
  real a_age[n_age];                                  // age varying coeffs (depart from linear fit)
  real a_eth[n_eth];                                  // ethnicity class varying coeffs 
  real a_income[n_income];                            // income varying coeffs (depart linear fit)
  real a_male[n_male];                                // gender varying coeffs
  real a_state[n_state];                              // state varying coeffs
  real<lower=0> sigma_age;                            // age var(sd) component
  real<lower=0> sigma_eth;                            // ethnicit var(sd) component
  real<lower=0> sigma_income;                         // income var(sd) component
  real<lower=0> sigma_male;                           // gender fluency var(sd) component
  real<lower=0> sigma_state;                          // state var(sd) component
}
// transformed parameters {                              // Model specification is here
//   vector[n] outcome_hat;
//   for (i in 1:n) {
//    outcome_hat[i] = b0 + a_age[age[i]] + a_eth[eth[i]] + a_income[income[i]] + a_male[male[i]] + a_state[state[i]];
//   }
// }
model {                                              // Model distributions for varying coefficients
  b0 ~ cauchy(0,2.5);                                // and prior distributions for unmodelled                                                                // Can modify Cauchy scale parameter
                                                     // Can also specify a uniform distribution here
  a_age ~ normal(0, sigma_age);                      // or by setting upper and lower limits for
  a_eth ~ normal(0, sigma_eth);                      // unmodelled parameters in the parameter step
  a_income ~ normal(0, sigma_income);                // above  
  a_male ~ normal(0, sigma_male);                      
  a_state ~ normal(0, sigma_state);

  sigma_age ~ cauchy(0,2.5);
  sigma_eth ~ cauchy(0,2.5);
  sigma_income ~ cauchy(0,2.5);
  sigma_male ~ cauchy(0,2.5);
  sigma_state ~ cauchy(0,2.5);

  for (i in 1:n) {
   outcome[i] = bernoulli_logit(b0 + a_age[age[i]] + a_eth[eth[i]] + a_income[income[i]] + a_male[male[i]] + a_state[state[i]]);
  }

//  outcome ~ bernoulli_logit(outcome_hat);            // For a continuous outcome, specify a
                                                     // normal distribution with an additional
                                                     // variance component, i.e.
                                                     // outcome ~ normal(outcome_hat, sigma_outcome)
                                                     // Note the prior distributions for 
                                                     // unmodelled parameters will change.
}
```

...THEN YOU CAN FIT.

```{r}
# Observations. 
downes_list <- list(
  outcome = sample$service_failure,
  male = sample$male + 1,
  age = sample$age, 
  eth = sample$eth, 
  income = sample$income,
  state = sample$state,
  n = length(sample$service_failure), 
  n_male = max(sample$male) + 1,
  n_age = max(sample$age),
  n_eth = max(sample$eth),
  n_income = max(sample$income),
  n_state = max(sample$state)
)
```

```{r, eval = FALSE}
fit <- stan(
  file = here::here("Code", "mrp", "downes_copy.stan"),
  data = downes_list,
  seed = 42 
)
```

### Stan user guide logit model

The data for the Stan User Guide model is specified with lower and upper bounds on the data for the number of categories for each variable. In this stan model, post-stratification takes place in the generated quantities section.

[Note to McKenna]
Questions to answer: 
What the epsilon, -epsilon is that?

```{stan stan-mrp, output.var="stan_mrp", eval=FALSE}
data {
  int<lower = 0> N;
  int<lower = 1, upper = 7> age[N];
  int<lower = 1, upper = 3> income[N];
  int<lower = 1, upper = 3> eth[N];
  int<lower = 1, upper = 2> male[N];
  int<lower = 1, upper = 50> state[N];
  int<lower = 0> y[N];
  int<lower = 0> P[7, 3, 3, 2, 50];
}
parameters {
  real alpha;
  real epsilon;
  real<lower = 0> sigma_beta;
  real<lower = 0> sigma_gamma;
  real<lower = 0> sigma_delta;
  real<lower = 0> sigma_rho;
  vector<multiplier = sigma_beta>[7] beta;
  vector<multiplier = sigma_gamma>[3] gamma;
  vector<multiplier = sigma_delta>[3] delta;
  vector<multiplier = sigma_rho>[50] rho;
}
model {
  y ~ bernoulli_logit(alpha + beta[age] + gamma[income] + delta[eth] + {epsilon, -epsilon}[male] + rho[state]);
  alpha ~ normal(0, 2);
  epsilon ~ normal(0, 2);
  sigma_beta ~ normal(0, 1);
  sigma_gamma  ~ normal(0, 1);
  sigma_delta ~ normal(0, 1);
  sigma_rho ~ normal(0, 1);
  beta ~ normal(0, sigma_beta);
  gamma ~ normal(0, sigma_gamma);
  delta ~ normal(0, sigma_delta);
  rho ~ norma(0, sigma_rho);
}
generated quantities {
  real expect_pos = 0;
  int total = 0;
  for (b in 1:7)
    for (c in 1:3)
      for (d in 1:3)
        for (f in 1:2)
          for (g in 1:50){
        total += P[b, c, d, f, g];
        expect_pos
          += P[b, c, d, f, g]
             * inv_logit(alpha + beta[b] + gamma[c] + delta[d] + {epsilon, -epsilon}[f] + rho[g]);
      }
  real<lower = 0, upper = 1> phi = expect_pos / total;
}

```

...THEN YOU CAN FIT.

```{r}
stan_data <- list( 
  age = sample$age, 
  eth = sample$eth, 
  income = sample$income, 
  male = sample$male, 
  state = sample$state, 
  Y = sample$service_failure, 
  N = length(sample$service_failure)
)
```

```{r, eval = FALSE}
fit <- stan(
  file = here::here("Code", "mrp", "stan_mrp.stan"),
  data = stan_data,
  seed = 42 
)
```

## Post-Stratification

Post-stratification consists of taking the sum of the estimates of the model for each cell times the number of people in each cell divided by the total number of people. Using post-stratificatio ncan improve accuracy of the estimates based on the information known.

IF YOU NEED TO INCLUDE THIS, WE WILL NEED SOMETHING IN THE YAML TO GET THIS WORK FOR A GITHUB_DOCUMENT (MARC WILL FIGURE THAT OUT). AN ALTERNATIVE IS TO EXPRESS AS CODE INSTEAD OF MATH.

THIS...

$$\omega=\displaystyle\frac{$$\sum_{j=1}^{J}\beta_{j}*N_{j}$$}{$$\sum_{j=1}^{J}N_{j}$$}$$

OR...

```
something / something
```

Post-stratification can be performed in stan using the generated quantities block.

```{r, eval = FALSE}
generated quantities {
  real expect_pos = 0;
  int total = 0;
  for (b in 1:7)
    for (c in 1:3)
      for (d in 1:3)
        for (f in 1:2)
          for (g in 1:50) {
        total += P[b, c, d, f, g];
        expect_pos
          += P[b, c, d, f, g]
             * inv_logit(b0 + a_age[b] + a_eth[c] + a_income[d] + a_male[f] + a_state[g]);
      }
  real<lower = 0, upper = 1> phi = expect_pos / total;
}           
```

## Compare to Population

Because we are using simulated data we can see how our post-stratified predictions compare to the true population.