### 2020-11-13 Code/mrp/sampledata.Rmd

---
title: "sampledata"
output: html_document
---

```{r setup, include=FALSE}
# Load packages.
library(tidyverse)
library(rstan)
library(bayesplot)
library(tidybayes)

# Set Stan options.
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

# Specify data and hyperparameter values.
sim_values <- list(
  N = 500,                            # Number of observations.
  K = 5,                              # Number of groups.
  I = 7,                              # Number of observation-level covariates.
  J = 3,                              # Number of population-level covariates.

  # Vector of group assignments.
  g = sample(5, 500, replace = TRUE),

  # Matrix of observation-level covariates.
  X = cbind(
    rep(1, 500),
    matrix(runif(500 * (7 - 1), min = 1, max = 10), nrow = 500)
  ),

  # Matrix of population-level covariates.
  Z = cbind(
    rep(1, 5),
    matrix(runif(5 * (3 - 1), min = 2, max = 5), nrow = 5)
  )
)

# Generate data.
sim_data <- stan(
  file = here::here("Code", "generate_data.stan"),
  data = sim_values,
  iter = 1,
  chains = 1,
  seed = 42,
  algorithm = "Fixed_param"
)

# Extract simulated data and parameters.
sim_y <- extract(sim_data)$y
sim_Gamma <- extract(sim_data)$Gamma
sim_Omega <- extract(sim_data)$Omega
sim_tau <- extract(sim_data)$tau
sim_Beta <- extract(sim_data)$Beta
sim_sigma <- extract(sim_data)$sigma

data <- list(
  N = sim_values$N,     # Number of observations.
  K = sim_values$K,     # Number of groups.
  I = sim_values$I,     # Number of observation-level covariates.
  J = sim_values$J,     # Number of population-level covariates.
  y = as.vector(sim_y), # Vector of observations.
  g = sim_values$g,     # Vector of group assignments.
  X = sim_values$X,     # Matrix of observation-level covariates.
  Z = sim_values$Z      # Matrix of population-level covariates.
)

fit_centered <- stan(
  file = here::here("Code", "hlm_centered.stan"),
  data = data,
  control = list(adapt_delta = 0.99),
  seed = 42
)